Efficiently stable presentations from error-correcting codes

### 2023-10-22 efficient_stability.tex

We begin by describing our construction. Given a graph $([k],E)$, the permutations $\{A_i\}_{i=1}^k$ act on  the set $\field^{[k]\cup E}$, namely bit strings  indexed by the vertices and edges of the graph. Now, each $A_i$ flips the $i^{\rm th}$ bit. Furthermore, it conditionally flips the  $ij^{\rm th}$  bit for $ij\in E$ where $i<j$, if the $j^{\rm th}$ bit is $1$. Namely, the $A_i$'s act as a NOT gate on the $i^{\rm th}$ bit composed with CNOT gates on all the $ij^{\rm th}$ bits, where $i<j$ and $ij\in E$. Now, all of these permutations are of order $2$. Equation \eqref{eq:commuting_or_not} is deduced by noting that  $A_iA_j=A_jA_i$ when $ij$ is not an edge, and that $A_iA_jA_iA_j$ has no fixed points when $ij$ is an edge --- this is because $A_iA_jA_iA_j$ must flip the $ij^{\rm th}$ bit.  
The von Neumann algebra $\mM$ containing the $A_i$'s is the one acting on $\complex^{\field^{[k]\cup E}}$ with its standard normalized trace $\tau^\cM(X)=\frac{1}{2^{k+|E|}}\Tr(X)$.
\begin{remark}
    A version of this construction (viewed in a different way) was used by Kozlov--Meshulam \cite{kozlov2019quantitative} to upper bound the Cheeger constant of the $k$-dimensional hypercube --- see Section 4.1 therein.
\end{remark}
\begin{claim}
    The map $x_i\mapsto A_i\in \cM$ is a $(\nicefrac{|E|}{\binom{k}{2}},\mu_R)$-approximate representation of the presentation \eqref{eq:z2-efficient} of $\Z_2^k$.
\end{claim}

@@

By combining Claims ??? we deduce Lemma \ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k}.

>>>>

We begin by describing our construction. Given a graph $([k],E)$, the permutations $\{A_i\}_{i=1}^k$ act on  the set $\field^{[k]\cup E}$, namely bit strings  indexed by the vertices and edges of the graph. Now, each $A_i$ flips the $i^{\rm th}$ bit. Furthermore, it conditionally flips the  $ij^{\rm th}$  bit for $ij\in E$ where $i<j$, if the $j^{\rm th}$ bit is $1$. Namely, the $A_i$'s act as a NOT gate on the $i^{\rm th}$ bit composed with CNOT gates on all the $ij^{\rm th}$ bits, where $i<j$ and $ij\in E$. Now, all of these permutations are of order $2$. Equation \eqref{eq:commuting_or_not} is deduced by noting that  $A_iA_j=A_jA_i$ when $ij$ is not an edge, and that $A_iA_jA_iA_j$ has no fixed points when $ij$ is an edge --- this is because $A_iA_jA_iA_j$ must flip the $ij^{\rm th}$ bit.  
The von Neumann algebra $\mM$ containing the $A_i$'s is the one acting on $\complex^{\field^{[k]\cup E}}$ with its standard normalized trace $\tau^\cM(X)=\frac{1}{2^{k+|E|}}\Tr(X)$.
\begin{remark}
    A version of this construction (viewed in a different way) was used by Kozlov--Meshulam \cite{kozlov2019quantitative} to upper bound the Cheeger constant of the $k$-dimensional hypercube --- see Section 4.1 therein.
\end{remark}
\begin{claim}\label{claim:appendix1}
    The map $x_i\mapsto A_i\in \cM$ is a $(\nicefrac{|E|}{\binom{k}{2}},\mu_R)$-approximate representation of the presentation \eqref{eq:z2-efficient} of $\Z_2^k$.
\end{claim}

@@

By combining Claims \ref{claim:appendix1} and \ref{claim:appendix2} applied to a graph which is a mathcing with $c$ edges, we deduce Lemma \ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k}.
